---
title: "viz_notebook"
author: "Zack Larsen"
date: "2/24/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(tidyverse, gridExtra, magrittr, rbokeh, plotly, ggthemes, ggbeeswarm, cowplot, ggalt, scales, ggcorrplot, ggExtra, RColorBrewer, randomcoloR, MASS,
d3heatmap, ggridges, r2d3, viridis, ggrepel, lubridate,
gapminder, superheat, ggthemr, networkD3)
```

## Data tables {.tabset}

### match.data

```{r, cache=TRUE, echo=TRUE}
match.data <- data.frame(varname = c("readmissions",
                                     "inpatient visits",
                                     "outpatient visits",
                                     "professional visits",
                                     "age"),
                         before = c(0.24,
                                    0.31,
                                    0.15,
                                    0.18,
                                    0.12),
                         after = c(0.17,
                                   0.20,
                                   0.07,
                                   0.11,
                                   0.15))

match.data
```

### Difference in differences

```{r, cache=TRUE, echo=FALSE}
diff_in_diff <- data.frame(varname = c("readmissions",
                                       "inpatient visits",
                                       "outpatient visits",
                                       "professional visits",
                                       "total net paid",
                                       "total spend",
                                       "BH total net paid",
                                       "BH total spend"),
                           difference = c(3.4,
                                          7.1,
                                          10.2,
                                          4.3,
                                          245,
                                          320,
                                          -112.134,
                                          -150))

diff_in_diff <- diff_in_diff %>% 
  mutate(sign = sign(difference),
         transform = sign*log(abs(difference))) %>% 
  arrange(desc(difference))

diff_in_diff
```

### mtcars

```{r}
mtcars
```

### diamonds

```{r}
diamonds
```

## Histograms {.tabset}

### ggplot2

```{r, cache=TRUE}
h1 <- ggplot(data = diamonds,
            mapping = aes(x = price))

h1 + geom_histogram(fill = "pink", color = "black", bins = 30) +
  labs(title="Pink histogram", y="Count of points in bin", x="Diamond price",
       caption="Data source: diamonds") + 
  theme(plot.title=element_text(size=30, 
                              face="bold", 
                              family="American Typewriter",
                              color="grey",
                              hjust=0.5,
                              lineheight=1.5),  # title
      plot.subtitle=element_text(size=15, 
                                 family="American Typewriter",
                                 face="bold",
                                 hjust=0.5),  # subtitle
      plot.caption=element_text(size=10),  # caption
      axis.title.x=element_text(vjust=-2,  
                                size=15),  # X axis title
      axis.title.y=element_text(size=15),  # Y axis title
      axis.text.x=element_text(size=10, 
                               angle=30,
                               vjust=0.5),  # X axis text
      axis.text.y=element_text(size=10))  # Y axis text



h3 <- ggplot(diamonds, aes(price, fill = cut(price, 100))) +
  geom_histogram(show.legend = FALSE, bins = 50) +
  theme_minimal() +
  labs(x = "Price", y = "n")

h3 + scale_fill_discrete(h = c(180, 360), c = 150, l = 80) +
  labs(title="Rad histogram") +
  theme(plot.background=element_rect(fill="salmon"), 
        plot.margin = unit(c(1, 2, 1, 1), "cm"),
        plot.title=element_text(size=30, 
                        face="bold", 
                        family="American Typewriter",
                        color="cyan",
                        hjust=0.5,
                        lineheight=1.5))



# Faceting with CUSTOM binwidth function
# Using density instead of count for a more even comparison across facets:
ggplot(data = diamonds, aes(x = price, stat(density))) +
  geom_histogram(fill = "pink", color = "black",
                 binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) +
  facet_wrap(~cut)
```

### rbokeh

```{r, cache=TRUE, warning=FALSE}
# 600 X 400 is a good figure to start with for the Viewer, but you
# might want more than that if you are going to zoom because it will
# not automatically adjust like ggplot
figure(width = 600, height = 400, legend_location = "top_right") %>%
  ly_hist(price, data = diamonds, breaks = 40, freq = FALSE) %>%
  ly_density(price, data = diamonds)

figure(width = 600, height = 400, legend_location = "top_right") %>%
  ly_hist(price, data = diamonds, breaks = 40, freq = FALSE, hover = price) %>%
  ly_density(price, data = diamonds)
```

### Plotly

```{r, cache=TRUE, warning=FALSE}
plot_ly(x = ~diamonds$price,
        type = "histogram",
        histnorm = "probability",
        alpha = 0.5)
```

## Density plots {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE}
# Proportionally scaled density map
d2 <- ggplot(data = diamonds,
             mapping = aes(x = price, fill = cut, color = cut))

d2 + geom_density(alpha = 0.3, mapping = aes(y = ..scaled..))
```

### rbokeh

```{r, cache=TRUE, warning=FALSE}
figure(width = 600, height = 400, legend_location = "top_right") %>%
  ly_hist(price, data = diamonds, breaks = 40, freq = FALSE, hover = (diamonds$price)) %>%
  ly_density(price, data = diamonds, color = "red")
```

## boxplots {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE}
x1 <- ggplot(data = diamonds,
             mapping = aes(x = cut, y = price)) +
  theme_economist() + 
  scale_fill_economist()

x1 + geom_boxplot(fill = "lightpink") + 
  coord_flip()
```

### rbokeh

```{r, cache=TRUE, warning=FALSE}
figure(ylab = "Height (inches)", width = 600) %>%
  ly_boxplot(voice.part, height, data = lattice::singer)

figure(data = lattice::singer) %>%
  ly_points(catjitter(voice.part), jitter(height), color = "black") %>%
  ly_boxplot(voice.part, height, with_outliers = FALSE)
```

### plotly

```{r, cache=TRUE, warning=FALSE}
p <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
p

p <- plot_ly(y = ~rnorm(50), type = "box") %>%
  add_trace(y = ~rnorm(50, 1))
p

p <- plot_ly(x = ~rnorm(50), type = "box") %>%
  add_trace(x = ~rnorm(50, 1))
p



p <- plot_ly(ggplot2::diamonds, y = ~price, color = ~cut, type = "box")
p
```

## violin plots {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE}
p <- ggplot(mtcars, aes(factor(cyl), mpg))
p + geom_violin() + geom_jitter(height = 0, width = 0.1)
p + geom_violin(aes(fill = cyl))



data(Boston)
dt.long <- gather(Boston, "variable",
                  "value", crim:medv)

col <- colorRampPalette(c("red", "blue"))(14)
# col.bp <- brewer.pal(9, "Set1") # brewer.pal only has a max of 9 colors
col.rc <- as.vector(distinctColorPalette(14))

ggplot(dt.long,aes(factor(variable), value))+
  geom_violin(aes(fill=factor(variable)))+
  geom_boxplot(alpha=0.3, color="black", width=.1)+
  labs(x = "", y = "")+
  theme_bw()+
  theme(legend.title = element_blank())+
  facet_wrap(~variable, scales="free")
```

### plotly

```{r, cache=TRUE, warning=FALSE}
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/violin_data.csv")

p <- df %>%
  plot_ly(
    x = ~day,
    y = ~total_bill,
    split = ~day,
    type = 'violin',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    )
  ) %>% 
  layout(
    xaxis = list(
      title = "Day"
    ),
    yaxis = list(
      title = "Total Bill",
      zeroline = F
    )
  )

p
```

## scatterplots {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE}

p <- ggplot(data = diamonds,
            mapping = aes(x = depth, y = price))

p + geom_point() +
  annotate(geom = "rect", xmin = 70, xmax = 80,
           ymin = 15000, ymax = 20000, fill = "red", alpha = 0.2) + 
  annotate(geom = "text", x = 70, y = 14000,
           label = "A surprisingly high \n recovery rate.", hjust = 0)
```

### rbokeh

```{r, cache=TRUE, warning=FALSE}
figure() %>%
  ly_points(Sepal.Length, Sepal.Width, data = iris, color = Species, hover = list(Sepal.Length, Sepal.Width))
```

## correlograms {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE}
corr <- round(cor(mtcars), 1)
# Plot
ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of mtcars", 
           ggtheme=theme_bw)
```

## bar charts {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE}
theme_set(theme_bw())  
# Data Prep
data("mtcars")  # load data
mtcars$`car name` <- rownames(mtcars)  # create new column for car names
mtcars$mpg_z <- round((mtcars$mpg - mean(mtcars$mpg))/sd(mtcars$mpg), 2)  # compute normalized mpg
mtcars$mpg_type <- ifelse(mtcars$mpg_z < 0, "below", "above")  # above / below avg flag
mtcars <- mtcars[order(mtcars$mpg_z), ]  # sort
mtcars$`car name` <- factor(mtcars$`car name`, levels = mtcars$`car name`)  # convert to factor to retain sorted order in plot.
# Diverging Barcharts
ggplot(mtcars, aes(x=`car name`, y=mpg_z, label=mpg_z)) + 
  geom_bar(stat='identity', aes(fill=mpg_type), width=.5)  +
  scale_fill_manual(name="Mileage", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(subtitle="Normalised mileage from 'mtcars'", 
       title= "Diverging Bars") + 
  coord_flip()
```

```{r, cache=TRUE, warning=FALSE}
avg_mpg <- mtcars %>%
        group_by(cyl, am) %>%
        summarise(mpg = round(mean(mpg, na.rm = TRUE),2))

p1 <- ggplot(avg_mpg, aes(factor(cyl), mpg, fill = factor(am))) +
        geom_bar(stat = "identity", position = "dodge", color = "grey40") +
        labs(fill = "AM group", caption = "Wide layout bar chart") +
        xlab("AM group") +
        ylab("# Cylinders") +
        theme_minimal() +
        ggtitle("MPG ratings with annotations") +
        scale_fill_brewer(palette = "Pastel1") +
        geom_text(aes(label = mpg), position = position_dodge(0.9),
                 vjust = 1.5, color = "white", family = "Georgia", size = 8)

p1
```

## principle components analysis {.tabset}

### base plot

```{r, cache=TRUE, warning=FALSE}
data(mtcars)
pcaCars <- princomp(mtcars, cor = TRUE)

# scree plot
plot(pcaCars, type = "l")

# cluster cars
carsHC <- hclust(dist(pcaCars$scores), method = "ward.D2")

# dendrogram
plot(carsHC)

# cut the dendrogram into 3 clusters
carsClusters <- cutree(carsHC, k = 3)

# add cluster to data frame of scores
carsDf <- data.frame(pcaCars$scores, "cluster" = factor(carsClusters))
carsDf <- transform(carsDf, cluster_name = paste("Cluster",carsClusters))
```

### ggplot2

```{r, cache=TRUE, warning=FALSE}
# ggplot2
p1 <- ggplot(carsDf,aes(x=Comp.1, y=Comp.2)) +
  theme_classic() +
  geom_hline(yintercept = 0, color = "gray70") +
  geom_vline(xintercept = 0, color = "gray70") +
  geom_point(aes(color = cluster), alpha = 0.55, size = 3) +
  xlab("PC1") +
  ylab("PC2") + 
  xlim(-5, 6) + 
  ggtitle("PCA Clusters from Hierarchical Clustering of Cars Data") 

# ggplot2 with ggrepel
p1 + geom_text_repel(aes(y = Comp.2 + 0.25, label = rownames(carsDf))) 
```

## ridgeline plots {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE, message=FALSE}
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016')
```

## hexbin density plots {.tabset}

### rbokeh

```{r, cache=TRUE, warning=FALSE, message=FALSE}
figure() %>% ly_hexbin(rnorm(10000), rnorm(10000))
```

## marginal histogram plots {.tabset}

### ggplot2

```{r, cache=TRUE, warning=FALSE, message=FALSE}
theme_set(theme_bw())  # pre-set the bw theme.
mpg_select <- mpg[mpg$hwy >= 35 & mpg$cty > 27, ]
g <- ggplot(mpg, aes(cty, hwy)) + 
  geom_count() + 
  geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="pink")
```

```{r, cache=TRUE, warning=FALSE, message=FALSE}
ggMarginal(g, type = "boxplot", fill="pink")
```
